---
  title: "Homework 6, Problem 1"
output: 
  html_notebook:
  toc: yes
toc_float: yes
author: Gabriel Johnson
---
  
rm(list=ls())

# Set working directory as needed
```{r}
library(dplyr)
library(rgl)
library(caret)
setwd("C:/Users/gabeb/Desktop/programming/CS 422 Homework/Gabriel,Johnson")
pts <- read.csv("perceptron.csv")
pts <- as.matrix(pts)

```
# The Perceptron function
#
# PARAMETERS:
# points: The dataset that is to be separated
# lamda:  The learning rate
# gamma:  The error threshold
#
# RETURNS
# A list containing three named elements.  The elements
# should be named: 
# "weights" - Contains the final weight vector (learned weights)
# "epochs"  - Number of epochs it took to converge
# "error"   - A vector of error calculated at the end of each epoch

```{r}
perceptron <- function(points, lamda, gamma) {
  w <- c(-gamma, runif(ncol(points) - 2))
  n <- nrow(points)
  label <- points[ , 1]
  obs <- points[ , 2:ncol(points)]
  misclassfied <- TRUE
  epochs <- 0
  error <- c()
  error2 <- c()
  mse = lamda * 2
  while (misclassfied) {
    mse = 0
    misclassfied <- FALSE
    for (i in 1:n) {
      if (label[i] * my_sign(obs[i , ], w) <= 0) {
        w <- w - lamda * (my_sign(obs[i , ], w) - label[i]) * obs[i , ]
        misclassfied <- TRUE
        err <- (((my_sign(obs[i,], w) - label[i])))
        error <- c(error, err)
      }
      epochs <- epochs + 1
      err2 <- sum(error)/n
      error2 <- c(error2,err2)
      error <- c()
    }
  }
  return(list(weights = w,epochs = epochs,error = error2))
}
```
# The sign function, this is the prediction function
# PARAMETERS:
# x : The X vector (input that needs to be predicted)
# weights: The weight vector
# RETURNS:
# -1 or 1
```{r}
my_sign <- function(x, weights) {
  net <- (sign(x %*% weights))
  return( net )
}
Plot2D <- function(points, a, b) {
  plot(points[, 3:4], xlab = "X", ylab = "Y",
       pch = ifelse(points[, 1] == 1, 2, 8),
       col = ifelse(points[, 1] == 1, "blue", "red"))
  abline(a, b)
}
```
# MAIN ENTRY POINT
```{r}
objlist <- perceptron(pts,0.5,0.01)
e <- unlist(objlist[1])
epochs <- unlist(objlist[2])
errors <- unlist(objlist[3])
Plot2D(pts, -e[1]/e[3], -e[2]/ e[3])
```
```{r}
plot((1:epochs),errors,type = "l",col = "red", xlab = "Epochs",ylab = "Error Rate", main = "Perceptron Training Error")
```



